The format of this lab follows the same format as the previous
ones. Your goal is to predict the value of the third column
(which will be missing on the test set) using the techniques
we have learned so far.

# Set up

Read in the following libraries and to load the nba dataset:

```{r}
library(readr)
library(dplyr)
library(ggplot2)
library(viridis)
library(e1071)
library(gam)

nba <- read_csv("https://statsmaths.github.io/ml_data/nba_shots.csv")
```

Make sure that your predictions are either `0` or `1`. Here are
some packages you might need:

```{r}
if (!require("e1071")) {
  install.packages("e1071")
}
```

# Lab 07

```{r}
#First, let us compute a model that include all the variables in an additive linear model. I decided to include a poly term for the shot clock, shot distance in an attempt to capture any non-linearity in the relationship between the variables. I also did the same for the closest defender distance and defender height. After playing around with the model parameters slightly, I found this to have the highest valid and train classification rate. 

model1 <- glm(fgm ~ poly(shot_clock, shot_dist, degree = 3) + touch_time + poly(close_def_dist ,defender_height, degree = 2) + period, data=nba, subset = train_id == "train", family = binomial(link="logit"))

model1_pred <- predict(model1, newdata = nba, type = "response")
model1_pred <- as.numeric(model1_pred > 0.5)

tapply(model1_pred == nba$fgm, nba$train_id, mean)

```

```{r}
#Next, I attempted to fit a similar model using SVM and the polynomial kernel to take into account the potential non-linear affects between variables.

y <- nba$fgm
X <- as.matrix(select(nba, shot_clock, shot_dist, touch_time, close_def_dist, defender_height ))

X_train <- X[nba$train_id == "train", ]
y_train <- y[nba$train_id == "train"]

#we see that using a polynomial kernel slightly improves our training set as well as our validation set. As a result, I have decided to go for model2. 

model2 <- svm(X_train, factor(y_train), kernel = "polynomial")
model2_pred <- as.numeric(as.character(predict(model2, newdata = X, type = "response")))
summary(model2)
tapply(model2_pred == nba$fgm, nba$train_id, mean)
```


# Submission

The code below assumes that you have added a prediction named
`fgm_pred` to every row of the dataset.

```{r}
nba$fgm_pred <- model2_pred
submit <- select(nba, obs_id, fgm_pred)
write_csv(submit, "class07_submit.csv")
```

Now, upload this file (ends with ".Rmd"), the HTML output
(ends with ".nb.html" or ".html"), and the csv file to
GitHub.
