The format of this lab follows the same format as the previous
ones. Your goal is to predict the value of the third column
(which will be missing on the test set) using the techniques
we have learned so far.

# Set up

Read in the following libraries and to load the crimes dataset:

```{r}
library(readr)
library(dplyr)
library(ggplot2)
library(viridis)
library(nnet)
library(FNN)

crimes <- read_csv("https://statsmaths.github.io/ml_data/chi_crimes_5.csv")
```

The names of the crimes are, in order:

```{r}
c("criminal_damage", "narcotics", "prostitution", "robbery", "theft")
```

Make sure that your predictions are either `1`, `2`, `3`, `4` or `5`.

# Lab 08
```{r}
#The first thing I will do is create a model matrix with the variables that may give some predictive power. I have decided to remove the location since our KNN will not work for categorial variables. I also decided to remove the day, month and hour variables since they reduced the overall train and valid set classification rates. 

X <- as.matrix(select(crimes, longitude, latitude))
y <- crimes$crime_type
X_train <- X[crimes$train_id == "train",]
y_train <- y[crimes$train_id == "train"]

#Next we run a preliminary KNN algorithm on our data, changing our k parameter to for tuning. To do this, I first started off with plugging in a few random values between 0 and 100 to gauge how smaller values of k affected the validation rate. I found that k values above 40 made the valid classification rate worse and worse. I looped over values from 10 to 40 and found that a k of 13 maximized the validation rate. 

#tuning_function <- function(){
  
 # result_final <- 0.33
 # i_final <- 0
  
 # for(i in 10:40){
 #   crimes$y_pred <- knn(train = X_train, test = X, cl = y_train, k = i)
  #  result <- tapply(crimes$y_pred == crimes$crime_type, crimes$train_id, mean)[3]
  #  if(result > result_final){
   #   result_final <- result
   #   i_final <- i
   # }
  #}
  
#  return(i_final)
#}


#result <- tuning_function()
#result
crimes$y_pred <- knn(train = X_train, test = X, cl = y_train, k = 13)
tapply(crimes$y_pred == crimes$crime_type, crimes$train_id, mean)



```


```{r}
#Next I try to add add the hour and month factors to the model. Our training and validation set were both higher after this change. 
model2 <- multinom(crime_type ~ crimes$y_pred + factor(year) + factor(hour) + factor(month) , data = crimes, subset = train_id == "train")

model2_pred <- predict(model2, newdata = crimes)
tapply(model2_pred == crimes$crime_type, crimes$train_id, mean)

table(y = crimes$crime_type, y_pred = model2_pred)

crimes$crime_type_pred <- model2_pred
```



# Submission

The code below assumes that you have added a prediction named
`crime_type_pred` to every row of the dataset.

```{r}
submit <- select(crimes, obs_id, crime_type_pred)
write_csv(submit, "class08_submit.csv")
```

Now, upload this file (ends with ".Rmd"), the HTML output
(ends with ".nb.html" or ".html"), and the csv file to
GitHub.
